Article data analysis

### 2018-10-30 notebooks/article_notebook.Rmd
added plot for mechanism per ST and species

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

@@

>>>>

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
always_allow_html: yes
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

@@

## Sequence types and resistance mechanisms
```{r, fig.width=14, fig.height=16}
# identify which sequence types are represented over three times
n_ST <- mlst_data[c("id","ST")] %>%
  group_by(ST) %>%
  count() %>%
  filter(n > 3)

# species ST  
species_st <- mlst_data[c("id","ST")] %>%
  left_join(isolate_data[c("id", "species")], by = "id") %>%
  select(-id) %>%
  group_by(ST, species) %>%
  count() %>%
  filter(ST %in% n_ST$ST)

# prepare micplot data for each ST with n > 3
micplot_st_df <- mlst_data[c("id", "ST")] %>%
  left_join(mic_data[c("id", "CIP")], by = "id") %>%
  group_by(ST, CIP) %>%
  count() %>%
  filter(ST %in% n_ST$ST)

# prepare heatmap data frame
st_heatmap_df <- mlst_data[c("id", "ST")] %>%
  left_join(mut_report[c("id", "marR")], by = "id") %>%
  left_join(cols_to_add, by = "id") %>%
  left_join(acquired_report, by = "id") %>%
  select(-id) %>%
  mutate_at(vars(-ST),
            funs(as.numeric)) %>%
  group_by(ST) %>%
  summarise_all(funs(sum, n = n())) %>%
  ungroup() %>%
  rename("n" = marR_n) %>%
  select(-contains("_n")) %>%
  gather(mechanism, value, -c(ST, n)) %>%
  mutate(mechanism = sub("_sum", "", mechanism),
         percent = round(value/n*100),
         value2 = ifelse(value > 0, 1, 0)) %>%
  filter(ST %in% n_ST$ST)

# plots
factor_breaks <- c(0.015,0.03,0.06,0.12,0.25,0.5,1,2,4,8,16,32,64,128,256)
factor_levels <- c("0.015","0.03","0.06","0.12","0.25","0.5","1","2","4","8","16","32","64","128","256")

## heatmap
heatmap_st <- ggplot(st_heatmap_df, aes(mechanism, ST, fill = percent))+
  geom_tile()+
  scale_fill_viridis()+
  guides(fill = FALSE)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3,
                                   size = 8),
        axis.title.x = element_blank())

## micplot
micplot_st <- ggplot(micplot_st_df, aes(ST, CIP))+
  geom_boxplot(fill = "#67a9cf")+
  scale_y_continuous(labels = factor_levels,
                     breaks = factor_breaks,
                     trans = "log10")+
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8),
        axis.ticks.y = element_blank())+
  coord_flip()

# n plot
nplot_st <- ggplot(species_st, aes(ST, n, fill = species))+
  geom_col(color = "black")+
  scale_fill_brewer(palette = "Paired")+
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 8),
        axis.ticks.y = element_blank(),
        legend.position = c(0.6, 0.95),
        legend.title = element_blank())+
  coord_flip()

cowplot::plot_grid(heatmap_st, micplot_st, nplot_st, ncol = 3, align = "h", rel_widths = c(3/6, 2/6, 1/6))
